\documentclass[11pt]{article}

\usepackage{amsmath, amssymb, mathrsfs}

\begin{document}

$$ O = \{ n  \mid  n = 2k + 1; k \in \mathbb{N} \} $$

\smallskip 

$$ \mathbb{R} $$

\section{sequences}

List of objects: order and repetition matter.
$$tap \neq tpa$$


\section{graphs}

Defined by sets of vertices and edges.

\smallskip

\noindent e.g., Friendships: what matters is
	
	who the people are (set of $V$ objects)
	
	who is friends with whom (set of $E$ edges)
	

\subsection{Types of Graphs:}
\textbf{affiliation graphs}

e.g., students and their courses

\noindent \textbf{conflict graphs}

e.g., courses with students in common conflict
	
	
\section{proofs.}
In the speed dating ritual, no one meets more than 12 people.

\textit{deductive proof}

In any round a person meets at most 3 new people.

There are 4 rounds, so one can only meet $3 * 4 = 12$ people total.

\smallskip

\subsection{when is a number a square}

\textit{even squares come from even numbers and even numbers have even squares.}

\smallskip

\textbf{proof.}

\begin{itemize}
\item $\bold{n\ even} \rightarrow n = 2k \rightarrow n^{2} = 2(2k^{2}) \rightarrow \bold{n^{2} even}$
\item $\bold{n\ odd} \rightarrow n = 2k + 1 \rightarrow n^{2} = 4(k^{2} + k) + 1 \rightarrow \bold{n^{2} odd}$ 
\end{itemize}

\subsection{graph}
\textbf{thm.} Any 6-node graph has a 3 person clique or a 3 person war.

\smallskip

\noindent \textit{proof.} in two cases.

\subsection{proof by contradiction.}
Ex. $\sqrt{2} $is irrational.

$$ \sqrt{2} = \Big\{ \frac{a_1}{b_1}, \frac{a_2}{b_2}, \frac{a_3}{b_3}, \cdots \Big\} $$

where $a_1, a_2, \dots$ are all integers and $b_1, b_2, \dots$ are all natural numbers.

Well ordering principle $\rightarrow$ there is a minimum $b_i$. Call it $b_*$.

$\sqrt{2} = \frac{a_*}{b_*}$, and $a_*$ and $b_*$ have no comment factor.

But this also implies that $a_*^{2} = 2b_*^{2}$, so $a_*$ is even.

And if $a_*$ is even, then $a_* = 2k \rightarrow b_*^{2} = 4k^{2}$, so $b_*$ is even.

This means, $a_*$ and $b_*$ have a common factor! But we know this to be false.

Thus $\sqrt{2}$ is irrational.



\section{definitions}

\textbf{axioms:} a self-evident statement that is asserted as true without proof.

\noindent \textbf{conjectures:} a claim that is believed true but is not true until proven so.

\noindent \textbf{theorems:} a proven truth. you can take it to the bank.




\end{document}